---
title: "ESM244 exploring death dataset"
author: "Danielle Bovenberg"
date: "5 March 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Supplementary data
For data per county, the census bureau makes them available here: 
https://www.census.gov/data/datasets/2016/demo/popest/counties-detail.html

##About this dataset
In the United States, police departments are not obligated to keep records about deaths. In a response to the increased relevance of these statistics, *The Guardian* started a crowdsourced dataset, inviting people to contribute information about known deaths due to police action. This dataset has been used by [authors].

While the dataset is in many ways 'the best we've got', it must be used with care. There are several facts that the dataset does *not* tell us. For example, it doesn't give insight into whether more deaths occurred in a given place, or whether more deaths were reported there. Second, since data is sourced from thousands of individuals, there may be differences in how variables have been interpreted, and thus entered. 


Futher questions: 
- What could be added to this dataset? 

Sources: 
Guardian site
Academic work


##Different stories to tell
The following options were explored; thinking of how best to navigate the dataset.



```{r subsets}

Death15_16$deaths <- rep(1,2226)
cityag <- aggregate(deaths ~ city, data = Death15_16, FUN = sum)
citysort <- arrange(cityag, -deaths)
citysort
citysort10 <- head(citysort, 10)
citysort10



tab2set <- Death15_16 %>% select(raceethnicity, gender, mannerofdeath, armed, year, season)

totaltable <- count(tab2set, raceethnicity)

tab2set_subset <- Death15_16 %>% filter(city == "Houston")
Houstontable <- count(tab2set_subset, raceethnicity) 

dta2 <- merge(oldtable, Houstontable, by="raceethnicity", all = T)


# New dataframe with top 10 cities
Death15_16$deaths <- rep(1,2226)
cityag <- aggregate(deaths ~ city, data = Death15_16, FUN = sum)
citysort <- arrange(cityag, -deaths)
citysort
citysort10 <- head(citysort, 10)
citysort10

# New dataframe with top 10 states
statelength<- aggregate(raceethnicity ~ state, data = Death15_16, FUN = length)
statesort <- arrange(stateag, -deaths)
statesort10 <- head(statesort, 10)
                         

A <- count(Death15_16, raceethnicity, state)
A2 <- arrange(A, state)
?count



```




```{r explore}

library(tidyverse)
library(mice)
library(ggplot2)
library(VIM)
library(leaflet)
library(scales)
library(ggrepel)
library(plotly)

Death15_16 <- read_csv("2015_2016.csv")
Death15_16$raceethnicity <- as.factor(Death15_16$raceethnicity)
Death15_16$gender <- as.factor(Death15_16$gender)
Death15_16$armed <- as.factor(Death15_16$armed)
Death15_16$age <- as.numeric(Death15_16$age)
Death15_16$month <- as.factor(Death15_16$month)
Death15_16$state <- as.factor(Death15_16$state)
Death15_16$mannerofdeath <- as.factor(Death15_16$mannerofdeath)

DF <- Death15_16
DF$city <- as.factor(DF$city)
DF$year <- as.factor(DF$year)
DF$gender <- as.factor(DF$gender)
DF$lawenforcementagency <- as.factor(DF$lawenforcementagency)

columbus <- DF %>% filter(city=="Columbus")
summary(columbus)

DF$monthno <- match(Death15_16$month, month.name)

DF$season <- ifelse(DF$monthno %in% 12 & 1:2, "winter",
                    ifelse(DF$monthno %in% 3:5, "spring",
                    ifelse(DF$monthno %in% 6:8, "summer",
                    ifelse(DF$monthno %in% 9:11, "fall", "other"))))

summary(DF)
ggplot(DF, aes(x = month)) +
  geom_bar()

DF_firearm <- DF %>% filter(armed == "Firearm")
DF_women <- DF %>% filter(gender == "Female")
ggplot(DF_women, aes(x = state)) +
  geom_bar()

DF_top3 <- DF %>% filter(city == "Los Angeles" | city == "Houston" | city == "Chicago")

ggplot(Death15_16, alpha = 0.2,
          aes(x = city, 
              group = gender,
              fill = gender)) + 
     geom_bar(position = "fill")


ggplot(DF_top3, alpha = 0.2, 
       aes(x = city, group = armed, fill = armed)) +
  geom_bar(position = "fill")

ggplot(DF_top3, alpha = 0.2, 
       aes(x = city, group = season, fill = season)) +
  geom_bar(position = "fill")

ggplot(DF_top3, 
       aes(x = city, group = raceethnicity, fill = raceethnicity)) +
  geom_bar(position = "fill")

ggplot(DF_top5states, 
       aes(x = state, group = season, fill = season)) +
  geom_bar(position = "fill")


DF_top5states <- DF %>%  filter(state == "CA" | state == "TX" | state =="FL"| state == "AZ" |state == "GA")
DF_top3states <- DF %>%  filter(state == "CA" | state == "TX" | state =="FL")

ggplot(DF_top3states, alpha = 0.2, 
       aes(x = state, group = armed, fill = armed)) +
  geom_bar(position = "fill")

ggplot(DF_top5states, alpha = 0.2, 
       aes(x = state, group = armed, fill = armed)) +
  geom_bar(position = "fill")


summary(DF_firearm)
ggplot(DF_firearm, aes(x = ))
```


```{r next}

# aggr(Death15_16, prop=F, numbers=T) 
# complete lat long with city, state

# Convert "month" variable to designate months with numbers
Death15_16$month <- match(Death15_16$month, month.name)
# Factors


#color coded

summary(Death15_16)

# Summary Counts


## Summary Statistics 2015-2016

#Race
summaryrace1516 <-  data.frame(table(Death15_16$raceethnicity)/length(Death15_16$raceethnicity)*100) 
colnames(summaryrace1516) <- c("Race/Ethnicity", "Percent")

#Average Age
summaryage1516 <- data.frame(mean(Death15_16$age, na.rm= TRUE))
colnames(summaryage1516) <- c("AverageAge")

#Month
summarymonth1516 <- data.frame(table(Death15_16$month)/length(Death15_16$month)*100) 
colnames(summarymonth1516) <- c("Month", "Percent")

#Armed/Not
summaryarmed1516 <- data.frame(table(Death15_16$armed)/length(Death15_16$armed)*100)
colnames(summaryarmed1516) <- c("Armed", "Percent")

#Gender
summarygender1516 <-data.frame(table(Death15_16$gender)/length(Death15_16$gender)) 
colnames(summarygender1516) <- c("Gender", "prop")

#plot_ly(summarygender1516, labels = ~Gender, values = ~prop, type = 'pie',
    #    textposition = 'auto',
   #     insidetextfont = list(color = "#FFFFFF"),

#marker = list(colors = c("#556677", "#AA3344", "#772200"),line = list(color = '#FFFFFF', width = 1)),
  #      textinfo='label+percent') %>%
  #layout(autosize=TRUE)

  #  title = "Title",
    #xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
   # yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)),
    #    insidetextfont = list(color = '#FFFFFF'),
 #marker = list(colors = c("#556677", "#AA3344", "#772200"),line = list(color = '#FFFFFF', width = 1))) 



# Maps of Percentages, etc

#label <- paste0(summarygender1516$Gender, " ", round(summarygender1516$prop *100), "%")
#ggplot(summarygender1516, aes(x="", y = prop, fill=Gender)) +
 # geom_bar(width=1, stat = "identity") +
  #coord_polar("y", start = 0) + 
  #theme_void() +
  #geom_text(aes(x = 1, y = cumsum(prop)-prop/2, label = label), size =3)

#testedit

```

